The Ultraviolet Radiation Environment of a Tropical Megacity in Transition: Mexico City 2000-2019

### 2021-07-02 Paper_third_version/Article_UV_Mexico.tex
Add changes from reviewer 3

The Mexico City Metropolitan Area is located at 19.4°N, 99.1°W, 2240
meters above sea level (asl), surrounded by mountain ridges exceeding
5000 m asl, with complex topography and thermal inversions that inhibit
winds and favor intense air pollution.\citep{Whiteman_2000,Fast_2007,Carre_n_Sierra_2015}
Air quality
monitoring and surface meteorological measurements in the MCMA are
conducted continuously by the Automated Atmospheric Monitoring Network
(RAMA, by its Spanish acronym) of the Atmospheric Monitoring System
(SIMAT, by its Spanish acronym) of the Mexico City government. Since the
year 2000, UV radiometers (model 501-A, Solar Light Company Inc.,
Glenside, PA) detecting wavelengths between 280-400 nm have been
measuring erythemally-weighted solar radiation. The calibration of the
UV sensors has been carried out annually by comparing against a
factory-calibrated reference sensor. The output voltages from the
network’s sensors have been compared for at least one week against the UV
readings from the reference sensor to derive a calibration factor. New
calibration factors typically differ from the old ones by 2\% or less.
Reference sensors have been also calibrated by the manufacturer and updated
periodically (between 3 to 5 years) to reduce systematic errors due to drifts in sensitivity.
Long term calibration drift are minimized by these yearly re-calibrations.
Although at the beginning only a few stations were in operation and have
been changing, currently 11 stations are recording erythemal
irradiances, which are then multiplied by 40 (see
Eq.~{\ref{eq:UVI}}) to give UV Indices.
Table~{\ref{table:stations}} describes the location of
the stations where the UV Index has been measured.
Figure~{\ref{fig:map}} shows how the radiometers of the
SIMAT have been distributed over MCMA, prioritizing sites with higher population density.
Near real-time data for each station are
available on the SIMAT official
website~\url{http://www.aire.cdmx.gob.mx/default.php}. Daily maximum
values (\(UVI_{\max}\)) were extracted~in each of the stations
around solar noon from the time interval from 11:00 h-15:00 h CST
(Central Standard Time). This database~with 7305 continuous days of
measurements during the period 2000-2019 was analyzed.

\begin{table}[H]
  \centering
  \begin{tabular}{ccccc}
    \hline
    Station  & Environment     & Lat (ºN) & Lon (ºW) & El (masl) \\ \hline
    CHO      & semi-urban      & 19.27    & 98.89    & 2253      \\
    CUT      & ecological park & 19.72    & 99.20    & 2263      \\
    FAC      & urban           & 19.48    & 99.24    & 2299      \\
    HAN      & urban           & 19.42    & 99.08    & 2235      \\
    LAA      & urban           & 19.48    & 99.15    & 2255      \\
    MER      & downtown        & 19.42    & 99.12    & 2245      \\
    MON      & rural           & 19.46    & 98.90    & 2252      \\
    MPA      & rural           & 19.18    & 98.99    & 2594      \\
    PED      & residential     & 19.33    & 99.20    & 2326      \\
    SAG      & urban           & 19.53    & 99.03    & 2241      \\
    SFE      & residential     & 19.36    & 99.26    & 2599      \\
    TLA      & urban           & 19.53    & 99.20    & 2311      \\
    UNAM$^*$ & University city & 19.33    & 99.18    & 2294      \\\hline
  \end{tabular}
  \caption{{{{SIMAT stations and AERONET site$^*$, environmental descriptors and geographical
                positions. Abbreviations names: Chalco (CHA), Cuautitlán (CUT), FES Acatlán
                (FAC), Hangares (HAN), Laboratorio de Análisis Ambiental (LAA), Merced
                (MER), Montecillo (MON), Milpa Alta (MPA), Pedregal (PED), San Agustín
                (SAG), Santa Fe (SFE), Tlalnepantla (TLA) and National Autonomous University of Mexico (UNAM$^*$)}}}}
  \label{table:stations}
\end{table}

@@

With the aim to explore the relationship between UVI and air pollutants
levels, the hourly averages for ozone (O\textsubscript{3}), carbon
monoxide (CO), nitrogen dioxide (NO\textsubscript{2}), sulfur
dioxide~(SO\textsubscript{2}) and particle matter with diameter
sizes~\(\le\) 10 ~$\mu$m (PM\textsubscript{10}), were downloaded
from the SIMAT.\cite{pollutants_data} For purposes of assessing the
influence on the UV Index at solar noon, only values obtained between
11h and 15h CST were considered for the trends analysis. Pollutant
measurements are conducted by the SIMAT using regulatory-grade
commercial instruments. Measurement principles include ultraviolet
photometry (model 400E, Teledyne-API) for O\textsubscript{3},
chemiluminescence (model 200E, Teledyne-API) for NO\textsubscript{2}, UV
fluorescence (model 100E, Teledyne API) for SO\textsubscript{2}, and
infrared absorption (model 300E, Teledyne-API) for CO. The
PM\textsubscript{10} continuous mass concentration was measured with
Tapered Element Oscillating Microbalance (TEOM 1400AB or TEOM 1405 DF,
Thermo Scientific) monitors. Gaseous pollutant levels are reported in
ppb concentration units for O\textsubscript{3}, SO\textsubscript{2} and
NO\textsubscript{2}, and in ppm for CO. Particulate matter mass
concentration is reported in $\mu$g m\textsuperscript{-3} at local
conditions for temperature and pressure.

@@

The UV Index data derived from satellite measurements of reflected radiance
were used for comparison with the ground-based measurements.
These data were provided by the Ozone Monitoring
Instrument (OMI)~on board of AURA-NASA satellite.\citep{dcio} OMI
is operated in cooperation between the Netherlands Agency for Aerospace
Programmes (NIVR), the Finnish Meteorological Institute (FMI) and NASA.
OMI performs observations over a geographical dimension of 13$\times$24km\textsuperscript{2} at nadir. For
the coordinates (19.33N, 99.18W) and elevation (2268 masl) of Mexico City,
the satellite overpass time is between 19:00h - 21:00h UTC (20:00h - 22:00h CET).
Measurements of the ozone profile and cloud cover are used via a
radiative transfer model to estimate the UVI at the ground.
The UV Index used for Mexico City corresponds to values at local noon time and clear sky
from OMUVB Level 2 data product, version 1.3 and collection 3.

@@

Calculations of the UV Index were also made with the Tropospheric
Ultraviolet Visible (TUV v5.3) model.\citep{Madronich_1987} The
model atmosphere was represented by 80 vertical layers, each 1 km thick,
starting at the 2.24 km asl elevation of MCMA, and for which the first
three km constitute the atmospheric boundary layer
(BL).\citep{Shaw_2007} The ozone profile above the BL is from the US
Standard Atmosphere, but rescaled to a value of 259.6 DU. Totals
including the BL contributions (of 13.7 DU in the year 2000 and 9.8 DU
in 2019, see below) were 273.1 DU in 2000 and 269.2 DU. The
climatological O\textsubscript{3} column for this latitude and season is
about 270 DU (plus or minus ca. 5 DU) so in good agreement with the
values used here.

@@

Above the atmospheric boundary layer, the model was taken to be free of
aerosols, NO\textsubscript{2}, or SO\textsubscript{2}. In one
sensitivity study, a total AOD of 0.7 was redistributed placing 0.2 in
the free troposphere (decreasing vertically with an exponential scale
height of 4 km, and 0.5 remaining in the BL). The calculated UVI
differed by less than 1\%. Thus, as long as the total AOD is known,
knowledge of the exact vertical aerosol profile is not critical towards
ground-level UVI -- but would obviously affect the vertical structure of
photolysis frequencies.

@@

Figure~{\ref{628947}} shows the diurnal variation of
the UVI for several specific cloud-free days, for different seasons and
several locations (CHO, MER, MON, PED, SAG, SFE and TLA). UV Index from
TUV model was used as reference of the behavior under clear sky days.
According to the comparison of measurements minute by minute, the dates
with prolonged fluctuations along the day were discarded. However,
during the rainy period (from June to October) at least a brief clouds
presence is common, as shown at CHO station around noon in 13 June
2017. Peak values range from 8 during autumn/winter to above 12
in spring/summer, in correspondence to the respective December and June
solstices. Although the stations are all within a 25 km radius,
substantial differences among them are notable. Survey of the locations
revealed that shadowing from nearby structures is not an issue. The good
agreement in the morning, followed by more divergence in the afternoon,
is consistent with the development of photo-chemical pollution hotspots
during the day. Previous studies (e.g., \citet{Castro_2001} and
\citet{Palancar_2013}) have shown
that surface UV radiation in Mexico City is attenuated significantly by
aerosols. The measurements shown in Fig.~{\ref{628947}}
are consistent with this increasing pollution during the course of the
day, with highest aerosol loading (and highest variability) attained in
the afternoon. Further support for the role of pollution in suppressing
the UVI comes from the observation made at the Santa Fe (SFE) site which
in Fig.~{\ref{628947}} are seen to be systematically
higher, e.g. by over 10\% in autumn afternoons, compared to the other
stations. The SFE station is displaced to the west from the majority of
the other stations, and remains in the outskirts. This site is also
approximately 300 m higher than Mexico City downtown, so that the
cleaner atmosphere results from both vertical and horizontal variations
in pollution levels.\citep{aire_cmdx_2017} It is indeed expected to have
higher values of the UVI, in agreement with the observations.

\begin{figure}[H]
  \begin{center}
    \includegraphics[width=0.70\columnwidth]{figures/season}
    \caption{{UV Index measured over MCMA by SIMAT stations each minute along the day
          under practically cloudless conditions for representative days of the
          year.
            {\label{628947}}%
        }}
  \end{center}
\end{figure}

The daily maximum UV Index of each station is denoted
by~\(UVI_{\max}\), all of them were counted in the period 2000-2019.
As shown in Figure~{\ref{461017}}, these values ranged
from 1 to 16, with a majority (61\%) of the days
experienced~\(UVI_{\max}\) values between 6 and 10, and remarkably
few, less than 1\%, in the higher 13-16 range.~The lowest values are
likely due to winter days with heavy cloud cover and low sun angles, and
UV attenuation by pollutants could also be amplified under such
conditions, due to longer photon path lengths at low sun and within
clouds. The extreme sparsity of high UVI values remains surprising, and
may be an indication of the rarity of extremely unpolluted days within
the city.

@@

Similar patterns are found when considering the monthly average of
the~\(UVI_{\max}\)~values as shown in
Figure~{\ref{310112}}. The long-term averages present a
seasonal variation (as in Fig.~{\ref{628947}}) that
follows approximately the cosine of the noontime solar zenith angle.
Notably, values rarely if ever exceed 12 (as in
Fig.~{\ref{461017}}). The lowest average UVI (near 7)
take place in winter while from March to August the values seem to be
flattened in the range 10-11. The rather low monthly UV Index values,
mainly could be a consequence of the presence of clouds in the rainy
season. However, urban aerosol pollution sources, biomass burning for
agriculture and wood cooking also contribute to poor air quality between
March-May~\citep{Retama_2015}. Using the
maximum~\(UVI_{\max}\) from all of stations every day (one daily
point), the monthly averages (\(\overline{UVI}_m\))
were calculated along the period 2000-2019 (except for June 2003 due to there were no measurements).
Long term trends in~\(\overline{UVI}_m\) are shown
in Figure~{\ref{185758}}. A clear~upward trend is seen,
with a slope for the linear fit of 0.9\%/year or +1.5 UVI units over the
two decades.

\begin{figure}[H]
  \begin{center}
    \includegraphics[width=0.70\columnwidth]{figures/Boxplotcos}
    \caption{{Boxplot of the monthly averages of the maximum UV Index values~(black
          dot) in MCMA for the period 2000-2019: median (central bold line),
          standard deviation (box edges),
          25\textsuperscript{th} and 75\textsuperscript{th} percentiles (the whiskers),
          the minimum and maximum  values (plus sign) and the cosine of the solar zenith angle at solar
          noon (blue curve).
            {\label{310112}}%
        }}
  \end{center}
\end{figure}

\begin{figure}[H]
  \begin{center}
    \includegraphics[width=0.70\columnwidth]{figures/UV_Moving_Average}
    \caption{{Moving average function (light gray curve) quarterly applied to monthly
          average UV Index, standard deviation of all data within a given month
          (black dots and dash line) and linear fit (red line).
            {\label{185758}}%
        }}
  \end{center}
\end{figure}

The UV Index computed from satellite-based observations
over the period 2005-2019 is mapped in
Figure~{\ref{485116}}. The satellite-derived UVIs vary
from 8 in winter to 16 in summer, both values being substantially higher
than the ground-based observations (ca. 7 for winter and 11 for summer,
see Fig.~{\ref{310112}}). We hypothesize that this
large difference between satellite-based estimation and ground-based
observation of the UV index is due to the intense air pollution of
Mexico City.~A rather similar behavior was detected in Santiago city,
Chile.\citep{Cabrera_2012}

Close inspection of Figure~{\ref{485116}} shows that the maximum values
are reached in June and July of each year.
While the early OMI estimates of the UVI did not account for the boundary layer aerosol absorption,
the current version (1.3) includes the postcorrection by \citet{Arola_2009}
for absorbing aerosols based on global monthly aerosol climatology.

@@

The observed changes in the concentrations of these air pollutants have
significant implications for surface UV radiation, as can be
demonstrated with the TUV radiative transfer model.
Table~{\ref{table:TUVmodel}} summarizes UVI values for
the June-July time period, estimated by the three methods: OMI
satellite-derived UVI, RAMA ground-based observations, and TUV modeling
using air pollution estimates. Two groups of values can be readily
identified: (1) RAMA daily record values, OMI with or without BL
aerosols, and TUV for very clean conditions, all with UVI values around
15-16; and (2) RAMA average daily maxima and TUV UVI using MCMA
pollutants as input, which are in good agreement for both 2018/19 and
2000/01 but much lower than the first group. Compared to the OMI
estimate that included the climatological aerosol correction (15.3),
observed RAMA values were lower by 35\% in 2000 and by 20\% in 2019.
Similarly, TUV values were 35\% lower in 2000 and 22\% lower in 2019
relative to the TUV values of 15.6 for a pristine atmosphere.

\begin{table}[H]
  \centering
  \begin{tabular}{lc}
    \hline
    \textbf{Conditions for estimation in June-July}     & \textbf{UV Index} \\ \hline
    RAMA maximum reached in period 2000-2019            & 15.0              \\ \hline
    RAMA average maxima 2018, 2019                      & 12.3              \\ \hline
    RAMA average maxima 2000, 2001                      & 9.9               \\ \hline
    OMI clear, local noon, no corrected for BL aerosols & 16.6              \\ \hline
    \begin{tabular}[c]{@{}l@{}}TUV "zero" pollution \\ AOD 0, O\textsubscript{3} 0 ppb, NO\textsubscript{2} 0 ppb, SO\textsubscript{2} 0 ppb\end{tabular}                          & 16.1              \\ \hline
    \begin{tabular}[c]{@{}l@{}
      }TUV pristine pollution \\ AOD 0.05, O\textsubscript{3} 10 ppb, NO\textsubscript{2} 0 ppb, SO\textsubscript{2} 0 ppb\end{tabular}                          & 15.6              \\ \hline
    \begin{tabular}[c]{@{}l@{}}TUV 2019:\\ AOD 0.5, O\textsubscript{3} 50 ppb, NO\textsubscript{2} 20 ppb, SO\textsubscript{2} 1 ppb\end{tabular}                          & 12.1              \\ \hline
    \begin{tabular}[c]{@{}l@{}}TUV 2000:\\ AOD 0.7, O\textsubscript{3} 70 ppb, NO\textsubscript{2} 40 ppb, SO\textsubscript{2} 20 ppb\end{tabular}                          & 10.2              \\ \hline
  \end{tabular}
  \caption{{UV Index monthly maximum estimates for June-July.}}
  \label{table:TUVmodel}
\end{table}

The agreement between RAMA observations and TUV model estimates is
excellent but also probably a bit fortuitous. Clouds on average reduce
the irradiance impingent on the surface, but scattering from them can
also cause transient enhancements (especially if the direct sunbeam is
not blocked) that could be recorded as daily maxima -- with cancellation
between these cloud effects resulting in improved agreement with the
cloud-free model. We cannot exclude that some of the observed trend in
UVI is due to changes in cloud cover. However, the modeled fractional
UVI reductions due to pollutants, shown in Table
  {\ref{table:TUVmodel}}, are in such good agreement with
the observed UVI reductions, that a compelling case can be made for a
dominant role of air pollutants in the long-term UVI trends.

Table~{\ref{table:year2000-2019}} shows the
contributions to UVI reductions from individual pollutants. Aerosols are
seen to be the major factor in both time periods, followed by
O\textsubscript{3}, NO\textsubscript{2}, and SO\textsubscript{2}. The
2000-2019 UVI increase is seen to result in comparable proportions from
fewer aerosols, less SO\textsubscript{2}, and the combined reductions in
O\textsubscript{3} and NO\textsubscript{2}.

@@

UV reductions by air pollutants are expected to be most severe near the
surface, while chemical reactions leading to photochemical smog occur
through the vertical extent of the BL.
Figure~{\ref{973680}} shows the vertical profiles of
photolysis coefficients (the reciprocals of photolytic lifetimes) for
two key reactions, the photolysis of O\textsubscript{3} to yield excited
oxygen atoms O(1D), and the photolysis of NO\textsubscript{2}. In the
absence of optically active pollutants, these coefficients would be
nearly independent of altitude in the BL. However, the presence of
pollutants leads to a strong decrease toward the surface, with notably
more severe reductions in 2000 compared to 2019.~While the specific
values shown in the figure are only illustrative for typical conditions,
routine daily air quality modeling should carefully account for the long
term variations in these photolysis coefficients.

>>>>

The Mexico City Metropolitan Area is located at 19.4°N, 99.1°W, 2240
meters above sea level (m asl), surrounded by mountain ridges exceeding
5000 m asl, with complex topography and thermal inversions that inhibit
winds and favor intense air pollution.\citep{Whiteman_2000,Fast_2007,Carre_n_Sierra_2015}
Air quality
monitoring and surface meteorological measurements in the MCMA are
conducted continuously by the Automated Atmospheric Monitoring Network
(RAMA, by its Spanish acronym) of the Atmospheric Monitoring System
(SIMAT, by its Spanish acronym) of the Mexico City government. Since the
year 2000, UV radiometers (model 501-A, Solar Light Company Inc.,
Glenside, PA) detecting wavelengths between 280-400 nm have been
measuring erythemally-weighted solar radiation. The calibration of the
UV sensors has been carried out annually by comparing against a
factory-calibrated reference sensor. The output voltages from the
network’s sensors have been compared for at least one week against the UV
readings from the reference sensor to derive a calibration factor. New
calibration factors typically differ from the old ones by 2\% or less.
Reference sensors have been also calibrated by the manufacturer and updated
periodically (between 3 to 5 years) to reduce systematic errors due to drifts in sensitivity.
Long term calibration drift are minimized by these yearly re-calibrations.
Although at the beginning only a few stations were in operation and have
been changing, currently 11 stations are recording erythemal
irradiances, which are then multiplied by 40 (see
Eq.~{\ref{eq:UVI}}) to give UV Indices.
Table~{\ref{table:stations}} describes the location of
the stations where the UV Index has been measured.
Figure~{\ref{fig:map}} shows how the radiometers of the
SIMAT have been distributed over MCMA, prioritizing sites with higher population density.
Near real-time data for each station are
available on the SIMAT official
website~\url{http://www.aire.cdmx.gob.mx/default.php}. Daily maximum
values (\(UVI_{\max}\)) were extracted~in each of the stations
around solar noon from the time interval from 11:00 h-15:00 h CST
(Central Standard Time). This database~with 7305 continuous days of
measurements during the period 2000-2019 was analyzed.

\begin{table}[H]
  \centering
  \begin{tabular}{ccccc}
    \hline
    Station  & Environment     & Lat (°N) & Lon (°W) & El (m asl) \\ \hline
    CHO      & semi-urban      & 19.27    & 98.89    & 2253       \\
    CUT      & ecological park & 19.72    & 99.20    & 2263       \\
    FAC      & urban           & 19.48    & 99.24    & 2299       \\
    HAN      & urban           & 19.42    & 99.08    & 2235       \\
    LAA      & urban           & 19.48    & 99.15    & 2255       \\
    MER      & downtown        & 19.42    & 99.12    & 2245       \\
    MON      & rural           & 19.46    & 98.90    & 2252       \\
    MPA      & rural           & 19.18    & 98.99    & 2594       \\
    PED      & residential     & 19.33    & 99.20    & 2326       \\
    SAG      & urban           & 19.53    & 99.03    & 2241       \\
    SFE      & residential     & 19.36    & 99.26    & 2599       \\
    TLA      & urban           & 19.53    & 99.20    & 2311       \\
    UNAM$^*$ & University city & 19.33    & 99.18    & 2294       \\\hline
  \end{tabular}
  \caption{{{{SIMAT stations and AERONET site$^*$, environmental descriptors and geographical
                positions. Abbreviations names: Chalco (CHA), Cuautitlán (CUT), FES Acatlán
                (FAC), Hangares (HAN), Laboratorio de Análisis Ambiental (LAA), Merced
                (MER), Montecillo (MON), Milpa Alta (MPA), Pedregal (PED), San Agustín
                (SAG), Santa Fe (SFE), Tlalnepantla (TLA) and National Autonomous University of Mexico (UNAM$^*$)}}}}
  \label{table:stations}
\end{table}

@@

With the aim to explore the relationship between UVI and air pollutants
levels, the hourly averages for ozone (O\textsubscript{3}), carbon
monoxide (CO), nitrogen dioxide (NO\textsubscript{2}), sulfur
dioxide~(SO\textsubscript{2}) and particle matter with diameter
sizes~\(\le\) 10 ~$\mu$m (PM\textsubscript{10}), were downloaded
from the SIMAT.\cite{atmosfrico} For purposes of assessing the
influence on the UV Index at solar noon, only values obtained between
11h and 15h CST were considered for the trends analysis. Pollutant
measurements are conducted by the SIMAT using regulatory-grade
commercial instruments. Measurement principles include ultraviolet
photometry (model 400E, Teledyne-API) for O\textsubscript{3},
chemiluminescence (model 200E, Teledyne-API) for NO\textsubscript{2}, UV
fluorescence (model 100E, Teledyne API) for SO\textsubscript{2}, and
infrared absorption (model 300E, Teledyne-API) for CO. The
PM\textsubscript{10} continuous mass concentration was measured with
Tapered Element Oscillating Microbalance (TEOM 1400AB or TEOM 1405 DF,
Thermo Scientific) monitors. Gaseous pollutant levels are reported in
ppb concentration units for O\textsubscript{3}, SO\textsubscript{2} and
NO\textsubscript{2}, and in ppm for CO. Particulate matter mass
concentration is reported in $\mu$g m\textsuperscript{-3} at local
conditions for temperature and pressure.

@@

The UV Index data derived from satellite measurements of reflected radiance
were used for comparison with the ground-based measurements.
These data were provided by the Ozone Monitoring
Instrument (OMI)~on board of AURA-NASA satellite.\citep{dcio} OMI
is operated in cooperation between the Netherlands Agency for Aerospace
Programmes (NIVR), the Finnish Meteorological Institute (FMI) and NASA.
OMI performs observations over a geographical area of 13$\times$24km\textsuperscript{2} at nadir. For
the coordinates (19.33°N, 99.18°W) and elevation (2268 m asl) of Mexico City,
the satellite overpass time is between 19:00h - 21:00h UTC (20:00h - 22:00h CET).
Measurements of the ozone profile and the attenuation by clouds are used via a
radiative transfer model to estimate the UVI at the ground.\cite{Tanskanen_2006}
The UV Index used for Mexico City corresponds to values at local noon time and clear sky
from OMUVB Level 2 OVP data product, version 1.3 and collection 3.

@@

Calculations of the UV Index were also made with the Tropospheric
Ultraviolet Visible (TUV v5.3) model.\citep{Madronich_1987} The
model atmosphere is represented by 80 vertical layers, each 1 km thick,
starting at the 2.24 km asl elevation of MCMA, and for which the first
three km constitute the atmospheric boundary layer
(BL).\citep{Shaw_2007} The ozone profile above the BL is from the US
Standard Atmosphere, but rescaled to a value of 259.6 DU. Total ozone columns
including the BL contributions (of 13.7 DU in the year 2000 and 9.8 DU
in 2019, see below) were 273.1 DU in 2000 and 269.2 DU in 2019, respectively. The
climatological O\textsubscript{3} column for this latitude and season is
about 270 DU (plus or minus ca. 5 DU) so in good agreement with the
values used here.

@@

Above the atmospheric boundary layer, the model atmosphere was specified to be,
NO\textsubscript{2}, or SO\textsubscript{2}. In one
sensitivity study, a total AOD of 0.7 was redistributed placing 0.2 in
the free troposphere (decreasing vertically with an exponential scale
height of 4 km, and 0.5 remaining in the BL). The calculated UVI
differed by less than 1\%. Thus, as long as the total AOD is known,
knowledge of the exact vertical aerosol profile is not critical for
ground-level UVI but would obviously affect the vertical structure of
photolysis frequencies.

@@

Figure~{\ref{628947}} shows the diurnal variation of
the UVI for several specific cloud-free days, for different seasons and
several locations (CHO, MER, MON, PED, SAG, SFE and TLA). UV Index calculated with
the TUV model was used as reference of the behavior during clear sky days.
According to the comparison of measurements minute by minute, the dates
with prolonged fluctuations along the day were discarded. However,
during the rainy period (from June to October) at least a brief episode of clouds,
as shown at CHO station around noon in 13 June
2017. Peak values range from 8 during autumn/winter to above 12
in spring/summer, in correspondence to the respective December and June
solstices. Although the stations are all within a 25 km radius,
substantial differences among them are notable. Survey of the locations
revealed that shadowing from nearby structures is not an issue. The good
agreement in the morning, followed by more divergence in the afternoon,
is consistent with the development of photo-chemical pollution hotspots
during the day. Previous studies (e.g., \citet{Castro_2001} and
\citet{Palancar_2013}) have shown
that surface UV radiation in Mexico City is attenuated significantly by
aerosols. The measurements shown in Fig.~{\ref{628947}}
are consistent with this increasing pollution during the course of the
day, with highest aerosol loading (and highest variability) attained in
the afternoon. Further support for the role of pollution in suppressing
the UVI comes from the observation made at the Santa Fe (SFE) site which
in Fig.~{\ref{628947}} are seen to be systematically
higher, e.g. by over 10\% in autumn afternoons, compared to the other
stations. The SFE station is displaced to the west from the majority of
the other stations, and remains in the outskirts. This site is also
approximately 300 m higher than Mexico City downtown, so that the cleaner atmosphere at SFE
is a result of the station’s location farther away from pollution sources,
both horizontally and vertically.\citep{SEDEMA2018a} It is indeed expected to have
higher values of the UVI, in agreement with the observations.

\begin{figure}[H]
  \begin{center}
    \includegraphics[width=0.70\columnwidth]{figures/season}
    \caption{{UV Index measurements at 1-minute resolution from several SIMAT stations within the MCMA
          under mostly cloudless conditions for representative days of the year.
            {\label{628947}}%
        }}
  \end{center}
\end{figure}

The daily maximum UV Index of each station is denoted
by~\(UVI_{\max}\), these values were counted in the period 2000-2019.
As shown in Figure~{\ref{461017}}, these values ranged
from 1 to 16, with a majority (61\%) of the days
experienced~\(UVI_{\max}\) values between 6 and 10, and remarkably
few, less than 1\%, in the 13-16 range.~The lowest values are
likely due to winter days with heavy cloud cover and low sun angles, and
UV attenuation by pollutants could also be amplified under such
conditions, due to longer photon path lengths at low sun and within
clouds. The extreme sparsity of high UVI values remains surprising, and
may be an indication of the rarity of extremely unpolluted days within
the city.

@@

Similar patterns are found when considering the monthly average of
the~\(UVI_{\max}\)~values as shown in
Figure~{\ref{310112}}. The long-term averages present a
seasonal variation (as in Fig.~{\ref{628947}}) that
follows approximately the cosine of the noontime solar zenith angle.
Notably, values rarely if ever exceed 12 (as in
Fig.~{\ref{461017}}). The lowest average UVI (near 7)
occurs in winter while from March to August the values seem to be
flattened in the range 10-11. The rather low monthly UV Index values
could be a consequence of the presence of clouds in the rainy
season. However, urban aerosol pollution sources, biomass burning for
agriculture and wood cooking also contribute to poor air quality between
March-May~\citep{Retama_2015}. Using the
maximum~\(UVI_{\max}\) from all of stations every day (one daily
point), the monthly averages (\(\overline{UVI}_m\))
were calculated along the period 2000-2019 (except for June 2003 due to there were no measurements).
Long term trends in~\(\overline{UVI}_m\) are shown
in Figure~{\ref{185758}}. A clear~upward trend is seen,
with a slope for the linear fit of 0.9\%/year or +1.5 UVI units over the
two decades.

\begin{figure}[H]
  \begin{center}
    \includegraphics[width=0.70\columnwidth]{figures/Boxplotcos}
    \caption{{Boxplot of the monthly averages of the maximum UV Index values~(black
          dot) in MCMA for the period 2000-2019: median (central bold line),
          interquartile range (box edges), 10\% and 90\% percentiles (the whiskers),
          the minimum and maximum values (asterisk) and the cosine of the solar zenith angle at solar
          noon (blue curve).
            {\label{310112}}%
        }}
  \end{center}
\end{figure}

\begin{figure}[H]
  \begin{center}
    \includegraphics[width=0.70\columnwidth]{figures/UV_Moving_Average}
    \caption{{3-month running average (violet segments curve) applied to monthly
          average UV Index (black line), the standard deviation (blue fill area) and linear fit (red line).
            {\label{185758}}%
        }}
  \end{center}
\end{figure}

The UV Index computed from satellite-based observations
over the period 2005-2019 is shown in
Figure~{\ref{485116}}. The satellite-derived UVIs vary
from 8 in winter to 16 in summer, both values being substantially higher
than the ground-based observations (ca. 7 for winter and 11 for summer,
see Fig.~{\ref{310112}}). We hypothesize that this
large difference between satellite-based estimation and ground-based
observation of the UV index is due to the intense air pollution of
Mexico City.~A rather similar behavior was detected in Santiago city,
Chile.\citep{Cabrera_2012}

Close inspection of Figure~{\ref{485116}} shows that the maximum values
are reached in June and July of each year.
While early OMI estimates of the UVI did not account for the boundary layer aerosol absorption,
the current version (1.3) includes the postcorrection by \citet{Arola_2009}
for absorbing aerosols based on global monthly aerosol climatology.

@@

The observed changes in the concentrations of these air pollutants have
significant implications for surface UV radiation, as can be
demonstrated with the TUV radiative transfer model.
Table~{\ref{table:TUVmodel}} summarizes UVI values for
the June-July time period, estimated by the three methods: OMI
satellite-derived UVI, SIMAT ground-based observations, and TUV modeling
using air pollution estimates. Two groups of values can be readily
identified: (1) SIMAT daily record values, OMI with or without BL
aerosols, and TUV for very clean conditions, all with UVI values around
15-16; and (2) the average of daily maxima of SIMAT observations and TUV UVI results using
MCMA pollutants as input. Note that SIMAT and TUV values are in good agreement
for both 2018-2019 and 2000-2001 and are considerably lower than for case (1). Compared to the OMI
estimate that included the climatological aerosol correction (15.3),
observed SIMAT values were lower by 35\% in 2000 and by 20\% in 2019.
Similarly, TUV values were 35\% lower in 2000 and 22\% lower in 2019
relative to the TUV values of 15.6 for a pristine atmosphere.

\begin{table}[H]
  \centering
  \begin{tabular}{lc}
    \hline
    \textbf{Conditions for estimation in June-July} & \textbf{UV Index} \\ \hline
    SIMAT average maxima 2000-2001                  & 9.9  $\pm$  1.2   \\ \hline
    SIMAT average maxima 2018-2019                  & 12.3 $\pm$  1.3   \\ \hline
    SIMAT maximum value reached in period 2000-2019 & 15.0              \\ \hline
    \begin{tabular}[c]{@{}l@{}}OMI maximum value reached in period 2005-2019,\\ not corrected for BL aerosols\end{tabular}                      & 16.6              \\ \hline
    \begin{tabular}[c]{@{}l@{}}TUV "zero" pollution \\ AOD 0, O3 0 ppb, NO2 0 ppb, SO2 0 ppb\end{tabular}                      & 16.1              \\ \hline
    \begin{tabular}[c]{@{}l@{}}TUV pristine pollution\\ AOD 0.05, O3 10 ppb, NO2 0 ppb, SO2 0 ppb\end{tabular}                      & 15.6              \\ \hline
    \begin{tabular}[c]{@{}l@{}}TUV 2019:\\ AOD 0.5, O3 50 ppb, NO2 20 ppb, SO2 1 ppb\end{tabular}                      & 12.1              \\ \hline
    \begin{tabular}[c]{@{}l@{}}TUV 2000:\\ AOD 0.7, O3 70 ppb, NO2 40 ppb, SO2 20 ppb\end{tabular}                      & 10.2              \\ \hline
  \end{tabular}
  \caption{{UVI maxima for June-July: averages and standard deviation from ground-based data
        for the years 2018-2019 and 2000-2001. UVI maximum value reached in the period 2000-2019 for
        UVI ground-based and the same for satellite in the period 2005-2019. TUV calculations for different conditions.
      }}
  \label{table:TUVmodel}
\end{table}

The agreement between SIMAT observations and TUV model estimates is
excellent but also probably a bit fortuitous. Clouds on average reduce
the irradiance impingent on the surface, but scattering from them can
also cause transient enhancements (these are only possible if the disk of the Sun is visible)
that could be recorded as daily maxima -- with cancellation
between these cloud effects resulting in improved agreement with the
cloud-free model. We cannot exclude that some of the observed trend in
UVI is due to changes in cloud cover. However, the modeled fractional
UVI reductions due to pollutants, shown in Table
  {\ref{table:TUVmodel}}, are in such good agreement with
the observed UVI reductions, that a compelling case can be made for a
dominant role of air pollutants in the long-term UVI trends.

Table~{\ref{table:year2000-2019}} shows the
contributions to UVI reductions from individual pollutants. Aerosols are
seen to be the major factor in both time periods, followed by
O\textsubscript{3}, NO\textsubscript{2}, and SO\textsubscript{2}. The
2000-2019 UVI increase results in comparable proportions from
less aerosols, less SO\textsubscript{2}, and the combined reductions in
O\textsubscript{3} and NO\textsubscript{2}.

@@

UV reductions by air pollutants are expected to be most severe near the
surface, while chemical reactions leading to photochemical smog occur
throughout the vertical extent of the BL.
Figure~{\ref{973680}} shows the vertical profiles of
photolysis coefficients (the reciprocals of photolytic lifetimes) for
two key reactions, the photolysis of O\textsubscript{3} to yield excited
oxygen atoms O(1D), and the photolysis of NO\textsubscript{2}. In the
absence of optically active pollutants, these coefficients would be
nearly independent of altitude in the BL. However, the presence of
pollutants leads to a strong decrease toward the surface, with notably
more severe reductions in 2000 compared to 2019.~While the specific
values shown in the figure are only illustrative for typical conditions,
routine daily air quality modeling should carefully account for the long
term variations in these photolysis coefficients.
